---
title: "COVID-19 Mapper: College Towns"
author: 
  - name: "Michelle Catherine Kelly"
    url: https://michelleckelly.github.io
date: "Last updated: `r format(Sys.Date())`"
output:
  html_document: 
    toc: true
    number_sections: true
    theme: paper
    highlight: tango
---

-----------

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(httr)
library(tidyverse)
library(lubridate)
library(gridExtra)
```

# Pull COVID-19 data

## Data availability statement

This data is compiled by the Bing COVID-19 Tracker effort (https://www.bing.com/covid) from multiple sources, including the [World Health Organization (WHO)](https://www.who.int/emergencies/diseases/novel-coronavirus-2019), [Centers for Disease Control and Prevention (CDC)](https://www.cdc.gov/coronavirus/2019-ncov/index.html), national and state public health departments, [BNO News](https://bnonews.com/index.php/2020/04/the-latest-coronavirus-cases/), [24/7 Wall St.](https://247wallst.com/), and [Wikipedia](https://en.wikipedia.org/wiki/2019%E2%80%9320_coronavirus_pandemic). The raw dataset can be accessed from the [GitHub Repo microsoft/Bing-COVID-19-Data](https://github.com/microsoft/Bing-COVID-19-Data). The dataset is updated daily at 3AM PST, and is released on a 24-hour delay.

## Code

```{r read data, echo=TRUE, warning = FALSE}
# Read data in from Bing Github repo
covid <- read_csv(url("https://media.githubusercontent.com/media/microsoft/Bing-COVID-19-Data/master/data/Bing-COVID19-Data.csv"), 
                  col_types = list("d", "c", "d", "d", "d", "d", "d", "d", "n",
                                   "n", "c", "c", "c", "c", "c"))

# Filter dataset to USA
USA <- covid %>% 
  filter(Country_Region == "United States") %>%
  mutate(Updated = mdy(Updated))

# Filter dataset to Michigan
MI <- covid %>% 
  filter(Country_Region == "United States" & AdminRegion1 == "Michigan") %>%
  mutate(Updated = mdy(Updated))

# Filter dataset to Houghton County (Michigan Tech)

# Additionally, calculate number of cases per 1,000 people
# Population of Houghton Co: 35,684 (source: googs, 2019)
# Population of Marquette Co: 66,699 (googs, 2019)

# Cases per 1,000 people = Total number of cases / (population size / 1,000)
HOTN <- covid %>% 
  filter(Country_Region == "United States" & AdminRegion1 == "Michigan" &
           AdminRegion2 == "Houghton County") %>%
  mutate(Updated = mdy(Updated),
         CasesPer1000 = Confirmed / (35684/1000),
         Confirmed_7dmean = zoo::rollmean(Confirmed, 7, fill = NA),
         ConfirmedChange_7dmean = zoo::rollmean(ConfirmedChange, 7, fill = NA),
         CasesPer1000_7dmean = zoo::rollmean(CasesPer1000, 7, fill = NA))

# Filter dataset to Marquette County (Northern MI Univ)
MQT <- covid %>% 
  filter(Country_Region == "United States" & AdminRegion1 == "Michigan" &
           AdminRegion2 == "Marquette County") %>%
  mutate(Updated = mdy(Updated), 
         CasesPer1000 = Confirmed / (66699/1000),
         Confirmed_7dmean = zoo::rollmean(Confirmed, 7, fill = NA),
         ConfirmedChange_7dmean = zoo::rollmean(ConfirmedChange, 7, fill = NA),
         CasesPer1000_7dmean = zoo::rollmean(CasesPer1000, 7, fill = NA))
```

# Plot

```{r plot, echo=T,warning=FALSE, fig.width=10, fig.height=8}
# since20201 <- TRUE subsets data to 1-Jan-2021 onward
since2021 <- TRUE

if(since2021) {
  # Subset only 2021 Houghton MI data
  HOTN <- HOTN %>% filter(Updated >= ymd("2021-01-01"))
  # Subset only 2021 MQT MI data
  MQT <- MQT %>% filter(Updated >= ymd("2021-01-01"))
  # Set frequency of date breaks
  dBreaks <- "1 week"
  } else {
  # If plotting all data, set to longer date breaks
  dBreaks <- "1 month"
  }


# Create unified palette
colors <- RColorBrewer::brewer.pal(name = "Set1", n = 3)
names(colors) <- c("Houghton", "Marquette")

# Plot: Cumulative cases
## Houghton Co
a <-
  ggplot(data = HOTN, aes(x = Updated)) +
    # Cases
    geom_bar(stat = "identity", aes(y = Confirmed, fill = "Houghton"), alpha = 0.2) +
    # 7 day average of cases
    geom_line(aes(y = Confirmed_7dmean, color = "Houghton")) +
    # Theme adjustments
    theme_classic() +
    scale_colour_manual(values = colors, limits = names(colors)) +
    scale_fill_manual(values = colors, limits = names(colors)) +
    scale_x_date(date_breaks = dBreaks) +
    labs(y = "Cumulative confirmed COVID-19 cases", 
         x = element_blank(), title = "Houghton County, Michigan",
         subtitle = "Population: 35,684") +
    theme(panel.border = element_rect(color = "black", fill = NA),
          axis.text = element_text(color = "black"), legend.position = "none",
          panel.grid.major.y = element_line(color = "grey95"),
          panel.grid.minor.y = element_line(color = "grey95"))

## Marquette Co
b <- 
  ggplot(data = MQT, aes(x = Updated, y = Confirmed)) +
    # Cases
    geom_bar(stat = "identity", aes(y = Confirmed, fill = "Marquette"), alpha = 0.2) +
    # 7 day average of cases
    geom_line(aes(y = Confirmed_7dmean, color = "Marquette")) +
    # Theme adjustments
    theme_classic() +
    scale_colour_manual(values = colors, limits = names(colors)) +
    scale_fill_manual(values = colors, limits = names(colors)) +
    scale_x_date(date_breaks = dBreaks) +
    labs(y = "Cumulative confirmed COVID-19 cases", 
         x = element_blank(), title = "Marquette County, Michigan",
         subtitle = "Population: 66,699") +
    theme(panel.border = element_rect(color = "black", fill = NA),
          axis.text = element_text(color = "black"), legend.position = "none",
          panel.grid.major.y = element_line(color = "grey95"),
          panel.grid.minor.y = element_line(color = "grey95"))

## Combined
c <- 
  ggplot(data = NULL, aes(y = CasesPer1000_7dmean)) +
    # 7 day average of cases per 1,000 people
    geom_line(data = HOTN, aes(x = Updated,  color = "Houghton")) +
    geom_line(data = MQT, aes(x = Updated, color = "Marquette")) +
    # Theme adjustments
    theme_classic() +
    scale_colour_manual(values = colors, limits = names(colors)) +
    scale_fill_manual(values = colors, limits = names(colors)) +
    scale_x_date(date_breaks = dBreaks) +
    labs(y = "Cumulative cases per 1,000 people", 
         x = element_blank()) +
    theme(panel.border = element_rect(color = "black", fill = NA),
          axis.text = element_text(color = "black"), legend.position = "none",
          panel.grid.major.y = element_line(color = "grey95"),
          panel.grid.minor.y = element_line(color = "grey95"))

# Plot: Confirmed cases per day
## Houghton Co
d <- 
  ggplot(data = HOTN, aes(x = Updated)) +
    # Cases
    geom_bar(stat = "identity", aes(y = ConfirmedChange, fill = "Houghton"), 
             alpha = 0.2) +
    # 7 day average of cases
    geom_line(aes(y = ConfirmedChange_7dmean, color = "Houghton")) +
    # Theme adjustments
    theme_classic() +
    scale_x_date(date_breaks = dBreaks) +
    scale_colour_manual(values = colors, limits = names(colors)) +
    scale_fill_manual(values = colors, limits = names(colors)) +
    labs(y = "New confirmed COVID-19 cases", 
         x = element_blank(), title = "Houghton County, Michigan",
         subtitle = "Population: 35,684") +
    theme(panel.border = element_rect(color = "black", fill = NA),
          axis.text = element_text(color = "black"), legend.position = "none",
          panel.grid.major.y = element_line(color = "grey95"),
          panel.grid.minor.y = element_line(color = "grey95"))

## Marquette Co
e <- 
  ggplot(data = MQT, aes(x = Updated)) +
    # Cases
    geom_bar(stat = "identity", aes(y = ConfirmedChange, fill = "Marquette"), alpha = 0.2) +
    # 7 day average of cases
    geom_line(aes(y = ConfirmedChange_7dmean, color = "Marquette")) +
    # Theme adjustments
    theme_classic() +
    scale_colour_manual(values = colors, limits = names(colors)) +
    scale_x_date(date_breaks = dBreaks) +
    scale_fill_manual(values = colors, limits = names(colors)) +
    labs(y = "New confirmed COVID-19 cases", 
         x = element_blank(), title = "Marquette County, Michigan",
         subtitle = "Population: 66,699") +
    theme(panel.border = element_rect(color = "black", fill = NA),
          axis.text = element_text(color = "black"), legend.position = "none",
          panel.grid.major.y = element_line(color = "grey95"),
          panel.grid.minor.y = element_line(color = "grey95"))

## Combined
f <- 
  ggplot(data = NULL, aes(y = ConfirmedChange_7dmean)) +
    # 7 day average of cases per 1,000 people
    geom_line(data = HOTN, aes(x = Updated,  color = "Houghton")) +
    geom_line(data = MQT, aes(x = Updated, color = "Marquette")) +
    # Theme adjustments
    theme_classic() +
    scale_colour_manual(values = colors, limits = names(colors)) +
    scale_fill_manual(values = colors, limits = names(colors)) +
    scale_x_date(date_breaks = dBreaks) +
    labs(y = "New confirmed COVID-19 cases", 
         x = element_blank()) +
    theme(panel.border = element_rect(color = "black", fill = NA),
          axis.text = element_text(color = "black"), legend.position = "none",
          panel.grid.major.y = element_line(color = "grey95"),
          panel.grid.minor.y = element_line(color = "grey95"))

# Arrange in grid
grob <- grid.arrange(a, b, c, d, e, f, nrow = 2)

# Save combined plot image to working directory
ggsave(grob, filename = "CovidMapper.png", device = "png", width = 10, 
       height = 8, units = "in")
```
